\chapter{Time metrology}


%====================================================================================================
\section{History and overview of metrologoy}

%----------------------------------------------------------------------------------------------------
\subsection{Metrology and reference values}

%----------------------------------------------------------------------------------------------------
\subsection{Time metrology through ages}
While precise time metrology is relatively young branch of measurement science humans were
measuring time from even before civilisation was formed.
In general it is possible to divide methods of time measurement into two categories based on 
what type of reference is used:
\begin{itemize}
	\item time of process
	\item count of occurences of periodic process
\end{itemize}

%----------------------------------------------------------------------------------------------------
\subsection{Modern approach with fous on GNSS related standards}
In modern day all measurement systems used in science and egineering are governed by international
standars. One of most well known set of standards are those provided by the International 
Organization for Standardization (IOS) \textbf{CITATION}.
In case of GNSS systems a standarized spatial and temporal frame must be decided on in order
for all element of system to work correctly.

\paragraph{Spatial reference frames}
In case of spatial information GNSS use two reference frames\textbf{CITATION}:
\begin{itemize}
	\item Earth-centered internal (ECI) for satellites
	\item Earth-centered Earth-fixed (ECEF) for ground recievers.
\end{itemize}
\textbf{SOME ECI RELATED IMAGE}
In case of satellites ECI is used as it is default reference frame for modelling spacecraft
motion \textbf{CITATION}. However in case of ground recievers that that are either fixed to 
specific point on earth or are moving from one geographical location to another ECEF is much
more natural way of describing position.
Reference frames for spatial data in GNSS systems are provief by:
\begin{itemize}
	\item International Terrestial Reference System (ITRS)
	\item International Celestial Reference System (ICRF)
\end{itemize}

\paragraph{Temporal reference frames}
When talking about time measurement reference frames in context of spaceborn objects a basic
concept from general relativity theory must be introduced.
One of those concept is \emph{proper time (PT)} which \textbf{WRITE THIS}.
Another is \emph{coordinate time (CT)} which denotes a perfect clock residing in a specific 
internal coordinate system \textbf{CITATION}.
In real life measurements CT is highest quality time measurement that we can achieve as a 
physical clock must reside in a physical space which by itself binds it to a coordinate system.
The highest precision time standad currently available is International Atomic Time (TAI)
that is based upon an ensemble of primary frequency standards locaed in multiple locations
around the globe. \textbf{CITATION} With TAI as a fequency standard the Coordinated Universal
Time (UTC) is calculated by adding a leap seconds for synchronization with earth rotation
period \textbf{CITATION}. UTC is now widely used standard for civilian timekeeping.

\paragraph{Temporal reference frames in GNSS systems}
While TAI provides highest precision it is not accessible in a real time and as such is 
not usefull for navigation purposes.
This is why each GNNS uses its own time refererence, in case of GPS a World Geodetic System
1984(WGS84) is used. In that standard timescales are provided by an aptly named GPS System Time
(GPST) which is realised by United States Naval Observatory (USNO). 
GPTS is used as a reference clock for both satellites and Earth components of GPS system and 
unlike UTC it is a continous timescale without leap seconds.
Relationship between GPTS and UTC can be described by an equation:
\begin{equation}
	b_{gps} = T_{utc} - T_{gps} = 19 + C_{0}.
\end{equation}
Where $C_{0}$ represents level of consistency between GPST and UTC.
\textbf{PLOT OF BIAS AND SOMETHING MORE ABOUT OTHER GNSS}
%====================================================================================================
\section{Mathematical clock model}

%----------------------------------------------------------------------------------------------------
\subsection{Discrete and continous clock models}
A mahematical model of a clock is a description of the bias in a sequence of time measurements.
Bias is a difference between the clock time, denoted as $T_{c}$, and the reference time,
denoted as a $T_{r}$. Both clock time and refeence time are going to be indexed with either
measurement number $i$ for a discrete model or a measurement timestamp  $t$ for a continous 
model. In case of continous model $T_{r}(t)=t$ as the reference clock is considered a
time source in model. Therefore bias can be described as a:
\begin{equation}
	b_{c}(i) = T_{r}(i) - T_{c}(i),
\end{equation}
for discrete model or a \textbf{FIX CONTINOUS EQUATION}
\begin{equation}
	b_{c}(t) = T_{r}(t) + \int_{t_0}^{t_1}T_{c}(\tau),
\end{equation}
for continous model.
In this work only discrete model in any furhter descriptions a term \emph{clock model} and
\emph{discreete clock model} will be used interchangably.



%----------------------------------------------------------------------------------------------------
\subsection{Basic clock model}

%----------------------------------------------------------------------------------------------------
\subsection{Oscilator noise model}

%----------------------------------------------------------------------------------------------------
\subsection{Standard clock model}


%====================================================================================================
\section{Stability analysis}

%----------------------------------------------------------------------------------------------------
\subsection{Allan variance}

%----------------------------------------------------------------------------------------------------
\subsection{Hadamard variance}

%----------------------------------------------------------------------------------------------------
\subsection{Sigma-tau curves}



%====================================================================================================
\section{Spaceborn clocks}

%----------------------------------------------------------------------------------------------------
\subsection{Clocks used in satellites}

%----------------------------------------------------------------------------------------------------
\subsection{Relativistic time transformation}

